---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# proxyconfig

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/proxyconfig)](https://cran.r-project.org/package=proxyconfig)


The goal of proxyconfig is to help setting the proxy interactively.  

## About proxy configuration

### Why you would need to configure a proxy ? 

In professional environment, there is often a proxy you need to pass through to escape your company network. When behind this firewall of your company, you need to first tell R that is needs to pass the proxy before trying to reach the url and not to use it when the url you want to access are inside the same network. 

### Examples of when it is needed

For example, behind a company proxy, 

+ `install.package` won't be able to reach the cran mirror that is outside. 
+ `download.file` won't be able to get a file from a url
+ `httr::GET`, `curl::curl` 

The first too are base R feature and have a method argument that you can use. Different method 
can have different settings. Most used is _libcurl_ that is also used by httr and curl :package:.

### Some words about proxy configuration in R

Sometimes, proxy is automatically picked up by R. It is the case on windows where R use the _wininet_ method by default, for which the 'Internet Options', used by Internet Explorer, are used for proxy configuration.   
On other system, proxy must be explicity configured. 

You'll find information on _Setting Proxies_ in R in the `utils::download.file()` help page:
`help("download.file", package = "utils")` or `?download.file`.


## Installation

Currently, this `r emo::ji("package")` is only available on [GitHub](https://github.com/) in development version :

```r
# install.packages("devtools")
devtools::install_github("cderv/proxyconfig")
```

or using

```r
source("https://install-github.me/cderv/proxyconfig")
```

## How to use `proxyconfig`

### How to set the proxy

Given a proxy url, you can set the proxy interactively using `set_proxy`. Proxy
url must be of the form _scheme://hostname[:port]_.

```{r set_proxy, eval = FALSE}
proxyconfig::set_proxy(proxy = "http://proxy.mycompany.com:3939")
```

By default, this will prompt the user for authentification ( _username_ and _password_ ). You can pass them in `set_proxy` argument too to use non-interactively. **This is not advice to pass them in clear in a script** - this is what the interactive mode with dialog box aims to prevent. 


If you don't have any authentification on your proxy, use empty values explicitly.
```{r set_proxy_without_auth, eval = FALSE}
proxyconfig::set_proxy(proxy = "http://proxy.mycompany.com:3939", username = "", password = "")
```

To prevent the proxy to be used for url on internal network domain, use `noproxy` argument. (empty by default). This useful for a github entreprise server or an internal cran repos for examples, respectively on `https://github.mycompany.com` and `https://cran.mycompany.com`. Both are on the same domain - `noproxy` is configured here to look for url on domain mycompany.com without exiting the internal pany network through the proxy.

```{r eval = FALSE}
proxyconfig::set_proxy(proxy = "http://proxy.mycompany.com:3939", noproxy = ".mycompany.com")
```

If several domains (or IP addresses) are necessary, they will be concatenated properly. 

```{r eval = FALSE}
proxyconfig::set_proxy(proxy = "http://proxy.mycompany.com:3939", noproxy = c(".mycompany.com", "163.104.50.180"))
```

```{r include = FALSE}
# for readme demo - not an example to follow as password is in clear
proxyconfig::set_proxy(proxy = "http://proxy.mycompany.com:3939", username = "its", password = "me", noproxy = c(".mycompany.com", "163.104.50.180"), https = TRUE)
```

### How to check what is configured ? 

`set_proxy()` will set the correct environment variable for the current session. You can verify if a proxy is currently configured with `is_proxy_activated()`

```{r check-proxy}
proxyconfig::is_proxy_activated()
```

You can have more information using `verbose = TRUE`. (Note that authentification is hidden when printed)

```{r check-proxy-verbose}
proxyconfig::is_proxy_activated(TRUE)
```

### What if a proxy is already set ? 

If a proxy is already set, `set_proxy` will issue a warning (and return false invisibly)
```{r}
(set_proxy(proxy =  "https://newproxy.company.com"))
```

You can unset a proxy configuration with `unset_proxy()`
```{r}
unset_proxy(verbose = TRUE)
# proxy is correctly deactivated
is_proxy_activated(TRUE)
```


# TODO

This is a very new package that works for me. Among the ideas I have for improvement

 - [ ] Use an option mechanism to make the proxy url persistent across sessions
 - [ ] Use [keyring](https://github.com/r-lib/keyring) to store and retrieve auth accross session
 - [ ] Make a templating system to use proxyconfig in an internal package. Inspiration `ghentr` and `pkgconfig`
 - [ ] Improve console output with `cli`
 - [ ] Create an add-on for better interactation (and keybinding)

